Something about cetacean heart rates and ballistocardiograms

### 2021-08-27 analysis/paper/paper.Rmd
Blue whale heart beat detection with jerk cue
Former-commit-id: fd01e15d2ec49284807899fcc372591105ff5c00

```{r bw-data}
fs <- 400
fny <- fs / 2
acc_filter <- signal::butter(4, c(1.5, 7.5) / fny, type = "pass") 
apply_filter <- function(x) signal::filtfilt(acc_filter, x)
window <- 1.8 #s
bw_400hz <- readRDS(here::here("analysis/data/derived_data/bw180905_53_400hz.rds")) %>% 
  mutate(across(c(surge, sway, heave), apply_filter, .names = "{.col}_filt"))

bw_10hz <- readRDS(here::here("analysis/data/derived_data/bw180905_53_10hz.rds")) %>% 
  mutate(jerk_smooth = tma(jerk, window * 10))

bw_elg <- readRDS(here::here("analysis/data/derived_data/bw180905_53_elg.rds"))
```

```{r bw-beats}
bw_400hz <- mutate(bw_400hz, surge_range = NA, bcg_beat = NA, bcg_bpm = NA)

for (row in seq(nrow(bw_eligible))) {
  i <- approx(bw_400hz$dt, 
              seq_along(bw_400hz$dt), 
              bw_elg[row, c("start", "stop")])$y
  i <- floor(i[1]):floor(i[2])
  region <- bw_400hz[i, ]
  surge <- region$surge_filt
  surge_range <- local_range(surge, window *fs)
  bcg_beat <- find_beats_lsr(surge, window, fs)
  bcg_bpm <- rep(NA, length(i))
  if (sum(bcg_beat) > 1) {
    bcg_dt <- region$dt[bcg_beat]
    bcg_bpm[bcg_beat] <- 1 / as.numeric(lead(bcg_dt) - bcg_dt, unit = "mins")
  }
  bw_400hz$surge_range[i] <- surge_range
  bw_400hz$bcg_beat[i] <- bcg_beat
  bw_400hz$bcg_bpm[i] <- bcg_bpm
}

@@

```{r bw-bcg}
ca_tz <- "Etc/GMT+7"
bcg_begin <- lubridate::ymd_hms("2018-09-05 21:29:05", tz = ca_tz)
bcg_end <- lubridate::ymd_hms("2018-09-05 21:30:30", tz = ca_tz)
bcg_sample <- filter(bw_400hz, between(dt, bcg_begin, bcg_end))
bcg_sample_10hz <- filter(bw_10hz, between(dt, bcg_begin, bcg_end))

@@

>>>>

```{r bw-data}
fs <- 400
fny <- fs / 2
acc_filter <- signal::butter(4, c(1.5, 7.5) / fny, type = "pass") 
apply_filter <- function(x) signal::filtfilt(acc_filter, x)
window <- 1.8 #s

bw_elg <- readRDS(here::here("analysis/data/derived_data/bw180905_53_elg.rds"))

bw_400hz <- readRDS(here::here("analysis/data/derived_data/bw180905_53_400hz.rds")) %>% 
  mutate(
    across(c(surge, sway, heave), apply_filter, .names = "{.col}_filt"),
    rid_left = approx(bw_elg$start, bw_elg$region_id, dt, "constant")$y,
    rid_right = approx(bw_elg$stop, bw_elg$region_id, dt, "constant", yleft = 0)$y + 1,
    region_id = ifelse(rid_left == rid_right, rid_left, NA)
  ) %>% 
  select(-c(rid_left, rid_right))

bw_10hz <- readRDS(here::here("analysis/data/derived_data/bw180905_53_10hz.rds")) %>% 
  mutate(jerk_smooth = tma(jerk, window * 10))
```

```{r bw-beats}
bw_400hz_beats <- bw_400hz %>% 
  mutate(jerk = approx(bw_10hz$dt, bw_10hz$jerk_smooth, bw_400hz$dt, rule = 2)$y) %>% 
  filter(!is.na(region_id)) %>%
  group_by(region_id) %>% 
  mutate(bcg_beat = find_beats_jerk(jerk, surge_filt, window, fs),
         bcg_bpm = bpm(bcg_beat, dt)) %>% 
  ungroup() %>% 
  select(dt, bcg_beat, bcg_bpm)

bw_400hz <- bw_400hz %>% 
  mutate(surge_range = local_range(surge_filt, window * fs)) %>% 
  left_join(bw_400hz_beats, by = "dt")

@@

```{r bw-bcg}
bw_bcg_plot <- function(rid) {
  ca_tz <- "Etc/GMT+7"
  bcg_sample <- filter(bw_400hz, region_id == rid)
  bcg_sample_10hz <- bw_10hz %>% 
    filter(between(dt, bcg_sample$dt[1], bcg_sample$dt[nrow(bcg_sample)]))

@@

  p <- cowplot::plot_grid(p1, p2, p3, p4, 
                          align = "v", 
                          axis = "lr",
                          nrow = 4, 
                          rel_heights = c(1.5, 0.75, 1, 1),
                          labels = "AUTO")

  ggsave(here::here(sprintf("analysis/figures/bw_bcg/bw_bcg_%02i.png", rid)), p)
}

foo <- lapply(unique(na.omit(bw_400hz$region_id)), bw_bcg_plot)
```
